---
title: "Introduction and EDA"
format: html
editor: visual
---

## Introduction

## Imports

```{r}
library(readr)
library(dplyr)
library(purrr)
library(ggplot2)
```

```{r}
diabetes_raw <- read_csv("./data/diabetes_binary_health_indicators_BRFSS2015.csv")
```

## Setting up the Data

I'd like to focus on lifestyle and demographic inputs (essentially asking whether I can predict diabetes without the medical markers), so we'll select and set up the following variables with better types or transformations where necessary.We also include BMI since it's the only numerical predictor and will make our analysis more interesting.

-   BMI

-   Smoker

-   PhysActivity

-   Fruits

-   Veggies

-   HvyAlcoholConsump

-   Education

-   Income

```{r}
spec(diabetes_raw)

predictors <- c("BMI", "Smoker", "PhysActivity", "Fruits", "Veggies", "HvyAlcoholConsump", "Education", "Income")

diabetes_raw <- diabetes_raw |> 
  select(all_of(c("Diabetes_binary", predictors)))

# no missing values, so we'll continue
diabetes_raw |>
  is.na() |>
  colSums()

diabetes_tbl <- diabetes_raw |> 
  mutate(BMI = as.numeric(BMI)) |> 
  mutate(across(all_of(c("Diabetes_binary", "Smoker", "PhysActivity", "Fruits", "Veggies", "HvyAlcoholConsump")), ~factor(.x, levels = c(0, 1), labels = c("No", "Yes")))) |> 
  mutate(Education = factor(Education, 
                            levels = c(seq(1:6)), 
                            labels = c(
                              "Never attended school or only kindergarten", 
                              "Grades 1 through 8 (Elementary)",
                              "Grades 9 through 11 (Some high school)",
                              "Grade 12 or GED (High school graduate)",
                              "College 1 year to 3 years (Some college or technical school)",
                              "College 4 years or more (College graduate)"))) |> 
  mutate(Income = factor(Income,
                         levels = c(1:8),
                         labels = c(
                           "Less than $10,000",
                           "$10,000 to less than $15,000",
                           "$15,000 to less than $20,000",
                           "$20,000 to less than $25,000",
                           "$25,000 to less than $35,000",
                           "$35,000 to less than $50,000",
                           "$50,000 to less than $75,000",
                           "$75,000 or more"
                         ))) |> 
  rename("DiabetesOutcome" = "Diabetes_binary")
```

## EDA

Let's check one-way counts and mean/sd of BMI.

```{r}
map(names(diabetes_tbl)[!(names(diabetes_tbl) %in% c("BMI"))], \(column_name) table(diabetes_tbl[column_name]))

mean(diabetes_tbl$BMI)
sd(diabetes_tbl$BMI)
```

Let's look for a couple of patterns, too. We'll get a sense of more ways this can fit together when we run models, but let's create a few two-way tables and look at how BMI fits into the picture.

#### Voluntary activities:

```{r}
temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$Smoker)
names(dimnames(temp)) <- c("Diabetes", "Smoker")
print(temp)

temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$HvyAlcoholConsump)
names(dimnames(temp)) <- c("Diabetes", "Heavy Alc. Use")
print(temp)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = Smoker), alpha = .5, bins = 45)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = HvyAlcoholConsump), alpha = .5, bins = 45)
```

Diabetes is notably more prevalent in this data among smokers than nonsmokers.

#### Food Intake:

```{r}
temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$Fruits)
names(dimnames(temp)) <- c("Diabetes", "Fruits")
print(temp)

temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$Veggies)
names(dimnames(temp)) <- c("Diabetes", "Veggies")
print(temp)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = Fruits), alpha = .5, bins = 45)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = Veggies), alpha = .5, bins = 45)
```

It looks like many people eat at least one serving of fruits and vegetables daily, which could make this a less powerful group of predictors.

#### Demographics

```{r}
temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$Education)
names(dimnames(temp)) <- c("Diabetes", "Education")
print(temp)

temp <- table(diabetes_tbl$DiabetesOutcome, diabetes_tbl$Income)
names(dimnames(temp)) <- c("Diabetes", "Income")
print(temp)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = Education), alpha = .5, bins = 20)

ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = Income), alpha = .5, bins = 20)
```

It's a bit harder to see patterns with more levels of each variable, but it does look like there is some impact between income level and diabetes in particular.

On the whole, there hasn't been a lot of appreciable difference in BMI histogram shape between levels of each variable. We'll do one last look at diabetes prevalence over a BMI histogram.

```{r}
ggplot(diabetes_tbl, aes(x = BMI)) +
  geom_histogram(aes(fill = DiabetesOutcome), alpha = .5, bins = 35)
```

It appears the "Yes" histogram could be slightly right-shifted compared to the "No," but that also could just be us reinforcing an expectation. Let's model to see what factors in most.